---
title: "Assignment3_code"
output: html_document
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup}
# Load required packages
library(tidyverse)
library(factoextra)
library(cluster)

# Load dataset
hotel_customers <- read.csv("/Users/andrejr/Documents/Universiteit/MSc Data Science & Marketing Analytics/Strategic Marketing Decision Making/Group assignment 3/HotelCustomersSubset_sample.csv")

# Quick look at the data
str(hotel_customers)
head(hotel_customers)
summary(hotel_customers)

# Set seed
set.seed(123)
```

```{r cleaning}
# Convert Age to integer
hotel_customers$Age <- as.numeric(hotel_customers$Age)

# Remove missing and unrealistic age values
hotel_customers <- hotel_customers %>%
  filter(!is.na(Age) & Age >= 18 & Age <= 95)
```

```{r select data include=FALSE}
# Select the variables
selected_variables <- c(
  "Age","AverageLeadTime","DaysSinceCreation",
  "LodgingRevenue","OtherRevenue",
  "PersonsNights","RoomNights",
  "DaysSinceLastStay","DaysSinceFirstStay"
)

# Keep columns of selected variables and check for missing values 
analysis_data <- hotel_customers %>%
  select(all_of(selected_variables)) %>%
  drop_na()
```

```{r scale include=FALSE}
# Standardize all variables
analysis_data_scaled <- scale(analysis_data)
```

```{r scale include=FALSE}
# Standardize all variables
analysis_data_scaled <- scale(analysis_data)
```

```{r optimal k include=FALSE}
# Check with elbow method
fviz_nbclust(analysis_data_scaled, kmeans, method = "wss", k.max = 13) 

# Check with silhouette method
fviz_nbclust(analysis_data_scaled, kmeans, method = "silhouette", k.max = 13) 
```

So in both cases k seems to be the optimal number of cluster
Elbow method: the curve starts to flatten after k = 4
Silhouette method: highest average silhouette width at k = 4

```{r kmeans include=FALSE}
kmeans_result <- kmeans(analysis_data_scaled, centers = 4, nstart = 25)
kmeans_result
```

- centers = 4, the found in Q1
- nstart = 25, → runs the algorithm 25 times with different starting points, after it keeps the best result -> this reduces the chance of getting stuck in a local minimum, as mentioned in Lecture 6 (slide 43)

