---
title: "Marketing Mix Modelling (MMM)"
subtitle: "Group Assignment 1"

---


```{r libraries}
# Core
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 4.5, dpi = 120
)

```
```

```{r read-data}
# Read data ---------------------------------------------------------------
stopifnot(file.exists("marketing_mix_assignment.csv"))
data <- readr::read_csv("marketing_mix_assignment.csv", show_col_types = FALSE)

```


```{r}

data <- data %>%
  rename(
    google_ads = `Google Ads`,
    facebook   = Facebook,
    tiktok     = TikTok,
    sales      = Sales,
    week       = Date
  )

# Now these work without backticks
colSums(is.na(select(data, google_ads, facebook, tiktok, sales)))

# Negative-value counts per column
sapply(select(data, google_ads, facebook, tiktok, sales), function(x) sum(x < 0, na.rm = TRUE))

# If you want separate vectors:
google_ads <- data$google_ads
facebook   <- data$facebook
tiktok     <- data$tiktok
sales      <- data$sales

```


```{r}
matplot(
  x = data$week,
  y = cbind(data$google_ads, data$facebook, data$tiktok),
  type = "l", lty = 1, lwd = 2,
  xlab = "", ylab = "Spend",
  main = "Ad Spend by Channel Over Time",
  col = c("steelblue", "dodgerblue3", "darkturquoise")
)
legend("topleft", inset = 0.02,
       legend = c("Google Ads", "Facebook", "TikTok"),
       col = c("steelblue", "dodgerblue3", "darkturquoise"),
       lty = 1, lwd = 2, bty = "n")




```












